---
title: "Assignment_5"
output: html_document
---

Below is the source data for the arrival status for 2 airlines
```{r echo=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
MyData <- read.csv(file="H:/CLASS/CUNY_SPS_MS_DS/Fall18_607/delayed_flights.csv", header=TRUE, sep=",")
MyData
```
For the purpose of this assignment this data must be made tidy. data is defined by tidy if it adheres to the following :

  Each variable forms a column.
  
  Each observation forms a row.
  
  Each type of observational unit forms a table.

This sample does  not follow this form. In order to organize it properly a variable needs to be added for count and the cities that are currently headers need to be made into observations.Before I could organize it, I needed to add headers for the airline and flight status columns, and remove the blank observation in the middle

```{r echo=FALSE}
MyData=na.omit(MyData)
names(MyData)[1]<-"Airline"
names(MyData)[2]<-"Status"
MyData
```
Then I used gather to restructure the variables and observations
```{r echo=FALSE}
restructured<-MyData %>%
  gather(city,count,-Airline,-Status)
restructured
```
Now the variable city is added and populated by what was previously headers. The count variable was added to represent the count of instances when a flight was on time or delayed for each city. Next I needed to fill in the blanks in the Airline column.These were actually filled with empty strings, in order to use the fill() function I needed to replace them with NA
```{r echo=FALSE}
restructured
is.na(restructured)<- restructured==''
restructured
filled<-restructured %>% fill(Airline)
filled
```
After that I grouped by airline and status to summarize the counts of flights into these categories
```{r echo=FALSE}
filled%>%group_by(Airline,Status)%>% summarise(B=sum(count))
```

