---
title: "NYC Subway Complaints"
author: "Anne Schwenker"
date: "December 8, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_Data,echo=FALSE,warning=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(tidyr)
library(plyr)
library(dplyr)
library(DATA606)
library(ggmap)
library(rvest)
library(ggplot2)
library(treemap)
library(data.table)
library(RSocrata)
library(sqldf)
library(leafletR)
library(tm)
```

#Problem Statement

In 2015 the average subway ridership had risen to its highest since 1948, only to drop by 2% by 2017 according the a NYT article "Subway Ridership Dropped Again in New York as Passengers Flee to Uber" that was released in August. https://nyti.ms/2n4nUMQ 

This drop in ridership is said to be due to poor service and poorly kept stations. Given the MTA has limited time and resources for repairs and improvements I intend to analyze publicly available data to create a combined rank for each of the 472 MTA subway stations in order to provide a list of subways prioritized by both high usage and high complaints and high incidents of 511 ev. 
---
#The Data
There is an abundance of publicly available data regarding ridership, station and entrance locations and customer complaints about subway stations. The sources I explored for this project are:
File |Area|Data Aquisition Method|Source
------------- | -------------
MTA_Customer_Feedback_Data__Beginning_2014 | Complaints | CSV hosted on github and accessed through URL |https://data.ny.gov/
511 NY MTA Events: Beginning 2010| Service Delays | CSV stored locally |https://data.ny.gov/
DOITT_SUBWAY_STATION_01_13SEPT2010| Locations | CSV hosted on github and accessed through URL | https://data.ny.gov/
NYC Transit Subway Entrance And Exit Data| Locations | API | https://data.ny.gov/
Turnstile_Usage_Data__2017.csv| Usage | CSV stored locally |https://data.ny.gov/
Subway Ridership Table| Usage | Screen Scraped |http://web.mta.info/nyct/facts/ridership/ridership_sub_annual.htm

Some of these sources were not appropriate for my purpose and were not used in the final analysis. 
An example of a source that was not ideal was using the turnstile dataset for usage, the turnstile dataset recorded cumulative ridership 4 times a day for every turnstile in the city. While usage could be found from this, it was simplier to access the MTA sourced figures from their website that were available at the station level. 
---

#Complaints Data
The Complaints Data is from the NYS open data, sourced from the MTA CRMS (Customer Relationship Management System) from 2014 to 2018. Becuase it was from a CRMS it was relatively clean and required very little clean up. I subsetted it for complaints, with the agency as subway where the line was not blank. Then I created a table from the line column to get counts and then gathered the records to collect counts of incident by subway line as the basis for the rank

```{r customer_complaints, echo=FALSE,warning=FALSE,message=FALSE}
#Read Customer Complaint Data 
customer_URL<-"https://raw.githubusercontent.com/aschwenker/607/master/MTA_Customer_Feedback_Data__Beginning_2014.csv"
customer_feedback <- read.csv(customer_URL,stringsAsFactors=FALSE)

#Subway Complaints Data Manipulation and Subsetting to Top 10
#Subset to complaints related to subways where line value isn't blank
subway_customer_feedback<-subset(customer_feedback,Commendation.or.Complaint == "Complaint" & Agency == "Subways" & Branch.Line.Route != 'No Value')
#Create table of counts and gather
subway_customer_feedback_line_table<-sort(table(subway_customer_feedback$Branch.Line.Route),decreasing=T)
line_table_df <-data.frame(rbind(subway_customer_feedback_line_table))

line_table_df_gathered<-line_table_df %>% gather(line,occurence)
#Remove leading X from numbered lines
gathered_clean<-c('X')
line_table_df_gathered$line<-gsub(paste0(gathered_clean,collapse = "|"),"", line_table_df_gathered$line)
#subset to top 10 by count of complaints related
line_top_10<-line_table_df_gathered %>% top_n(10)
#Plot top 10 by count descending order
p1<-ggplot(data=line_top_10, aes(x=reorder(line, -occurence), y=occurence, fill=line)) +
  geom_bar(position = 'dodge',stat="identity")+ theme(legend.position="none")+xlab(label="Subway Line")+
  ylab(label="Count of Related Complaints")+ggtitle("Top Ten Subways By Complaint Count")
p1
```
---
#511 Data
511 incident data is accessed from NYS open data portal and sourced from NYSDOT. each record is either an incident, accident, construction project, or special event. I subset this data for MTA Subway and MTA New York City Transit data. This dataset appears to have a lot more manual entry, because the facility_name field required a lot of clean up. Then I created a table from that column to get counts and then gathered the records to collect counts of incident by subway line as the basis for the rank
```{r 511_incidents, echo=FALSE,warning=FALSE,message=FALSE}
#511 reports
#read in 511 data from socrata api
Data_511<-read.socrata("https://data.ny.gov/resource/gvhy-7eph.csv")
head(Data_511)
#list organizations to determine subsetting
unique(Data_511$organization_name)
smaller <- subset(Data_511,organization_name == "MTA NYC Transit Subway" | organization_name == "MTA - New York City Transit")
unique(smaller$facility_name)
#Clean up names of subway lines
remove_words<-c("Epress",'Select',' es ','some','Linees','Lines','lines', 'Line','line','Trains','trains', 'Train','train', '  ','Tains', '#', ',','liine','and','&')
x <-smaller$facility_name        #Company column data
smaller$facility_name<- removeWords(x,remove_words)     #Remove stopwords
unique(smaller$facility_name)
smaller$facility_name<-gsub("&","",smaller$facility_name)
smaller$facility_name<-gsub("#","",smaller$facility_name)
smaller$facility_name<-gsub(",","",smaller$facility_name)

smaller<-smaller %>% 
  mutate(facility_name = strsplit(as.character(facility_name), " ")) %>% 
  unnest(facility_name)
head(smaller)
smaller<-smaller[!(is.na(smaller$facility_name) | smaller$facility_name==""), ]

Data_511_facility_name_table<-sort(table(smaller$facility_name),decreasing = T)
Data_511_facility_name_table
Data_511_facility_name_table <-data.frame(rbind(Data_511_facility_name_table))
Data_511_facility_name_table_gathered<-Data_511_facility_name_table%>% gather(line,occurence)
gathered_clean<-c('V','X')
Data_511_facility_name_table_gathered$line<-gsub(paste0(gathered_clean,collapse = "|"),"", Data_511_facility_name_table_gathered$line)
Data_511_facility_name_table_gathered_top_10<-Data_511_facility_name_table_gathered %>% top_n(10)
top_10_line_subset<-unique(Data_511_facility_name_table_gathered_top_10$line)

p511<-ggplot(data=Data_511_facility_name_table_gathered_top_10, aes(x=reorder(line, -occurence), y=occurence, fill=line)) +
  geom_bar(position = 'dodge',stat="identity")+ theme(legend.position="none")+xlab(label="Subway Line")+
  ylab(label="Count of Related 511 Incidents")+ggtitle("Top Ten Subways By 511 Incident Count")
p511
```

---
#Ridership Data
Ridership data was screen scraped from the MTA website. This data had figures for every year from 2012 to 2017 as well as ranks and change figures. This data was aggregated at the station level which meant it could not be combine with the customer complaints and 511 incident data
```{r ridership, echo=FALSE,warning=FALSE, message=FALSE}
#screen scrape
ridership_url<-("http://web.mta.info/nyct/facts/ridership/ridership_sub_annual.htm")
webpage <- read_html(ridership_url)
ridership <- ridership_url %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="subway"]') %>%
  html_table()
ridership<-ridership[1]
ridership <- data.frame(ridership)
ridership_subset<-c("Station..alphabetical.by.borough.","X2017.Rank")
ridership<-ridership[ridership_subset]
ridership<-ridership[!(is.na(ridership$X2017.Rank) | ridership$X2017.Rank==""), ]
ridership$X2017.Rank<-as.numeric(ridership$X2017.Rank)
ridership_sorted<-ridership[order(ridership$X2017.Rank),]
head(ridership_sorted)
```

---
#Complaint and Incident subject Matter
For both MTA customer complaints and 511 incident data I looked also at the explanation for the complaint or incident to see what the top causes were
```{r customer_complaints_plots, echo=FALSE,warning=FALSE,message=FALSE, error=FALSE}
#Revieiwing Top Complaint content
top_10_line_subset_customer<-unique(line_top_10$line)

customer_top_10<-subset(subway_customer_feedback,Branch.Line.Route %in% top_10_line_subset_customer & Subject.Detail != "No Value")

subway_customer_feedback_matter_table<-sort(table(customer_top_10$Subject.Matter),decreasing = T)
subway_customer_feedback_matter_table_df <-data.frame(rbind(subway_customer_feedback_matter_table))
subway_customer_feedback_matter_table_df_gathered<-subway_customer_feedback_matter_table_df%>% gather(subject_matter,occurence)
subway_customer_feedback_matter_table_df_gathered_top_10<-subway_customer_feedback_matter_table_df_gathered %>% top_n(10)
#Plot top ten events happening on top ten highest incident lines
pcustomer_subject_matter<-ggplot(data=subway_customer_feedback_matter_table_df_gathered_top_10, aes(x=reorder(subject_matter, -occurence), y=occurence, fill=subject_matter)) +
  geom_bar(position = 'dodge',stat="identity")+xlab(label="Subject Matter")+theme(legend.position="none")+
  ylab(label="Count of Related MTA Customer Complaints")+ggtitle("Top Ten MTA Customer Complaint Subject Matters Occuring on Top 10 highest incident subway lines By MTA Customer Complaint Count")+ coord_flip()
pcustomer_subject_matter

#Top events on these lines
smaller_top_10_event_type_table<-sort(table(smaller_top_10$event_type),decreasing = T)
smaller_top_10_event_type_table_df <-data.frame(rbind(smaller_top_10_event_type_table))
smaller_top_10_event_type_table_df_gathered<-smaller_top_10_event_type_table_df%>% gather(event_type,occurence)
smaller_top_10_events_top_10<-smaller_top_10_event_type_table_df_gathered %>% top_n(10)
#Plot top ten events happening on top ten highest incident lines
p511_events<-ggplot(data=smaller_top_10_events_top_10, aes(x=reorder(event_type, -occurence), y=occurence, fill=event_type)) +
  geom_bar(position = 'dodge',stat="identity")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+xlab(label="Event")+
  ylab(label="Count of Related 511 Incidents")+ggtitle("Top Ten 511 Events Occuring on Top 10 highest incident subway lines By 511 Incident Count")
p511_events

```
---
#Data Challenges

Different Data Sources had different ways of aggregating the information so it was difficult to create a combined rank. The 511 data and the customer complaints are both collected at the line level while usage is collected at the station level. Becuase of this I ended up ranking lines by 511 incidents and customer complaints and did not use ridership at all. 
---
#
```{r RANK, echo=FALSE}
#RANK
#restrict to top 10
customer_line_ranked<-rank(subway_customer_feedback_table[1:10],ties.method= "first")
customer_line_ranked_df <-data.frame(rbind(customer_line_ranked))
customer_line_ranked_df_gathered<-customer_line_ranked_df %>% gather(line,occurence)
#Remove leading X from numbered lines
gathered_clean<-c('X')
customer_line_ranked_df_gathered$line<-gsub(paste0(gathered_clean,collapse = "|"),"", customer_line_ranked_df_gathered$line)
customer_line_ranked_df_gathered
```

```{r pressure, echo=FALSE}
plot(pressure)
```


