Below is the source data for the arrival status for 2 airlines
```{r echo=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
MyData <- read.csv(file="H:/CLASS/CUNY_SPS_MS_DS/Fall18_607/delayed_flights.csv", header=TRUE, sep=",")
MyData
```
For the purpose of this assignment this data must be made tidy. data is defined by tidy if it adheres to the following :

  Each variable forms a column.
  
  Each observation forms a row.
  
  Each type of observational unit forms a table.

The three samples I chose were all from NYC Open data. 
OCDV Family Justice Center Foreign Born Report
https://data.cityofnewyork.us/Public-Safety/OCDV-Family-Justice-Center-Foreign-Born-Report/bray-s6dc

2017 Intimate Partner Violence Related Snapshots: New York City Community Board Districts
https://data.cityofnewyork.us/Public-Safety/2017-Intimate-Partner-Violence-Related-Snapshots-N/qiwj-eg47

New York City Population by Borough, 1950 - 2040
https://data.cityofnewyork.us/City-Government/New-York-City-Population-by-Borough-1950-2040/xywu-7bv9


for OCDV Family Justice Center Foreign Born Report, it had counts for survey question respodents where is survey question was a separate variable.
```{r echo=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
family_justice_foreign_Born <- read.csv(file="H:/CLASS/CUNY_SPS_MS_DS/Fall18_607/OCDV_Family_Justice_Center_Foreign_Born_Report.csv", header=TRUE, sep=",")
family_justice_foreign_Born

```
I used gather to restructure this into a tidy format so that there was one variable for survey question, and one for count, with an observation for each survey question and year
```{r echo=FALSE}

restructured<-family_justice_foreign_Born %>%
  gather(Intake_Question,response_count,-Visit_Year,)
restructured_clean <-restructured[!(is.na(restructured$response_count) | restructured$response_count==""), ]
restructured_clean
```
Then I summarized to find the number of respondants by year
```{r echo=FALSE}
restructured_clean%>%group_by(Visit_Year)%>% summarise(Total_Response_Count=sum(response_count))
```
For 2017 Intimate Partner Violence Related Snapshots: New York City Community Board Districts the data was structured to show for each community district the counts of incidences reported. Each incident was its own variable.It also had a totals row for one variable.
```{r echo=FALSE}
Intimate_Partner_Violence <- read.csv(file="H:/CLASS/CUNY_SPS_MS_DS/Fall18_607/2017_Intimate_Partner_Violence_Related_Snapshots__New_York_City_Community_Board_Districts.csv", header=TRUE, sep=",")
Intimate_Partner_Violence

```
First I removed any row where the community district or borough column was na or blank to remove the totals row
```{r echo=FALSE}
Intimate_Partner_Violence_clean <- Intimate_Partner_Violence[!(is.na(Intimate_Partner_Violence$Comm_Dist_.Boro) | Intimate_Partner_Violence$Comm_Dist_.Boro==""), ]
Intimate_Partner_Violence_clean <-Intimate_Partner_Violence_clean[!(is.na(Intimate_Partner_Violence_clean$Comm_District) | Intimate_Partner_Violence_clean$Comm_District==""), ]
```
Then I gathered it so that all incident types for each community district were an observation and the count was another variable
```{r echo=FALSE}
restruct_Intimate_Partner_Violence_clean <- Intimate_Partner_Violence_clean %>% 
  gather(incident, reporting_count,IPV_DIR:DV_Rape,-Comm_Dist_.Boro)
restruct_Intimate_Partner_Violence_clean
```

After that I rolled it up to count of all incidents by borough

```{r echo=FALSE}
restruct_Intimate_Partner_Violence_clean%>%group_by(Comm_Dist_.Boro)%>% summarise(Total_reporting_count=sum(reporting_count))

```
For New York City Population by Borough, 1950 - 2040 the data is organized with an observation for every borough and an NYC total then the variables are each year the population and percent.
```{r echo=FALSE}
NYC_Pop_By_Boro <-read.csv(file="H:/CLASS/CUNY_SPS_MS_DS/Fall18_607/New_York_City_Population_by_Borough__1950_-_2040.csv", header=TRUE, sep=",")
NYC_Pop_By_Boro
```
I wanted to structure it so that the year was a variable, the borough was a variable and the population value and percent were each variables.
In order to acheive this first I united population and percentage by using unite so that I could gather them with just one observation for each year
```{r echo=FALSE}
united<-NYC_Pop_By_Boro%>%
  unite(x1950,X1950,X1950...Boro.share.of.NYC.total,sep="_")%>%
  unite(x1960,X1960,X1960...Boro.share.of.NYC.total,sep="_")%>%
  unite(x1970,X1970,X1970...Boro.share.of.NYC.total,sep="_")%>%
  unite(x1980,X1980,X1980...Boro.share.of.NYC.total,sep="_")%>%
  unite(x1990,X1990,X1990...Boro.share.of.NYC.total,sep="_")%>%
  unite(x2000,X2000,X2000...Boro.share.of.NYC.total,sep="_")%>%
  unite(x2010,X2010,X2010...Boro.share.of.NYC.total,sep="_")%>%
  unite(x2020,X2020,X2020...Boro.share.of.NYC.total,sep="_")%>%
  unite(x2030,X2030,X2030...Boro.share.of.NYC.total,sep="_")%>%
  unite(x2040,X2040,X2040...Boro.share.of.NYC.total,sep="_")
united
```

Then I gathered by year and population 

```{r echo=FALSE}
restructured_NYC_Pop_By_Boro<-united%>%
  gather(year,population,-Age.Group,-Borough)
restructured_NYC_Pop_By_Boro
```
Then I broke out the united population and percent by using separate
```{r echo=FALSE}
separated<-restructured_NYC_Pop_By_Boro %>%
  separate(population,c("population","percent_total"),sep = "_")
separated
```

Finally I subsetted to view only 1950, 1980 and 2000 to see the change
```{r echo = false}
pop_1950<-subset(separated,year=="x1950"|year=="x1980"|year=="x2000",select=Borough:population)
pop_1950
```

